2.3: OLS Linear Regression

### 2019-09-25 static/homeworks/01-problem-set-answers.Rmd
Add problem set 1 answers, add broom logo in slides 2.3

*Due by Thursday, September 19, 2019*

- [<i class="fas fa-file-pdf"></i> `Download as PDF`](/homeworks/01-problem-set.pdf)
- [<i class="fab fa-markdown"></i> `04-r-practice-answers.Rmd`](/homeworks/01-r-problem-set.Rmd)

# Instructions

For this problem set, you may type up your answers to each question (save plots as images and include them with your answers) and also turn in a (commented!) *.R* file of commands for each relevant question. You may email or turn in a paper copy. It is not relevant for this particular problem set, but in the future, I will allow handwritten answers for math/theory questions.

Alternatively, you may download the `.Rmd` file, do the homework in markdown, and email to me a single `knit`ted `html` or `pdf` file (and be sure that it shows all of your code).

To minimize confusion, I suggest creating a new `R Project` (e.g. `hw1`) and storing any data and plots in that folder on your computer. See my [example workflow](https://github.com/ryansafner/workflow).

You may work together (and I highly encourage that) but you must turn in your own answers. I grade homeworks 70% for completion, and for the remaining 30%, pick one question to grade for accuracy - so it is best that you try every problem, even if you are unsure how to complete it accurately.

@@

```{r}
# save as a new tibble
top_5_boys_2017 <- babynames %>% # take data
  filter(sex=="M", # filter by males
         year==2017) %>% # and for 2007
  arrange(desc(n)) %>% # arrange in largest-to-smallest order of n (number)
  slice(1:5) %>% # optional, look only at first 5 rows
  mutate(percent = round(prop*100, 2)) # also optional, make a percent variable rounded to 2 decimals

@@

```{r}
# save as a new tibble
top_5_girls_2017 <- babynames %>% # take data
  filter(sex=="F", # filter by females
         year==2017) %>% # and for 2007
  arrange(desc(n)) %>% # arrange in largest-to-smallest order of n (number)
  slice(1:5) %>% # optional, look only at first 5 rows
  mutate(percent = round(prop*100, 2)) # also optional, make a percent variable rounded to 2 decimals

@@

```{r}
babynames %>%
  group_by(year) %>%
  filter(sex == "M",
         year>1980) %>%
  arrange(desc(n))%>%
  slice(1)
```

@@

```{r}
babynames %>%
  group_by(year) %>%
  filter(sex == "F",
         year>1980) %>%
  arrange(desc(n))%>%
  slice(1)
```

@@

top_boys_names<-c("James", "John", "Robert", "Michael", "William")
```

@@

top_girls_names<-c("Mary", "Elizabeth", "Patricia", "Jennifer", "Linda")
```

@@

```{r}
babynames %>%
  mutate(male = ifelse(sex == "M", n, 0),
         female = ifelse(sex == "F", n, 0)) %>%
  group_by(name) %>%
    summarize(Male = sum(male),
              Female = sum(female))%>%
  filter(Male!=0,
         Female!=0) %>%
  mutate(perc_male = (Male/(Male+Female)*100)) %>%
  arrange(perc_male) %>%
  filter(perc_male > 48,
         perc_male < 52) %>%
  mutate(total = Male+Female) %>%
  arrange(desc(total)) %>%
  slice(1:10)
```

@@

```{r}
# note these file paths will be different for you
polfreedom<-read.csv(here::here("./static/data/freedomhouse2018.csv"))
econfreedom<-read.csv(here::here("./static/data/econfreedom.csv"))

polfreedom

econfreedom
```

## 9. 
**The `polfreedom` dataset is still a bit messy. Let's overwrite it (or assign to something like `polfreedom2`) and select `Country.Territory` and `Total` (total freedom score) and rename `Country.Territory` to `Country`.**

```{r}
polfreedom<-polfreedom %>%
  select(Country.Territory, Total) %>%
  rename(Country=Country.Territory)
```

@@

@@

```{r, fig.retina=3}
# install.packages("ggrepel")
library(ggrepel)

>>>>

Answers may be longer than I would deem sufficient on an exam. Some might vary slightly based on points of interest, examples, or personal experience. These suggested answers are designed to give you both the answer and a short explanation of *why* it is the answer.

@@

```{r}
# save as a new tibble
top_5_boys_2017 <- babynames %>% # take data
  filter(sex=="M", # filter by males
         year==2017) %>% # and for 2007
  arrange(desc(n)) %>% # arrange in largest-to-smallest order of n (number)
  slice(1:5) %>% # optional, look only at first 5 rows; head(., n=5) also works
  mutate(percent = round(prop*100, 2)) # also optional, make a percent variable rounded to 2 decimals

@@

```{r}
# save as a new tibble
top_5_girls_2017 <- babynames %>% # take data
  filter(sex=="F", # filter by females
         year==2017) %>% # and for 2007
  arrange(desc(n)) %>% # arrange in largest-to-smallest order of n (number)
  slice(1:5) %>% # optional, look only at first 5 rows; head(., n=5) also works
  mutate(percent = round(prop*100, 2)) # also optional, make a percent variable rounded to 2 decimals

@@

```{r}
babynames %>%
  group_by(year) %>%
  filter(sex == "M",
         year>1980) %>%
  arrange(desc(n))%>%
  slice(1) # take first row only
```

@@

```{r}
babynames %>%
  group_by(year) %>%
  filter(sex == "F",
         year>1980) %>%
  arrange(desc(n))%>%
  slice(1) # take first row only
```

@@

# make a vector of the names (we'll need this for our graph below)
top_boys_names<-c("James", "John", "Robert", "Michael", "William")
```

@@

# make a vector of the names (we'll need this for our graph below)
top_girls_names<-c("Mary", "Elizabeth", "Patricia", "Jennifer", "Linda")
```

@@

There's a lot to this, so I'll break this up step by step and show you what happens at each major step.

We want to find the names where 48% to 52% of the babies with the name are male, as I defined in the footnote. First let's `mutate` a variable to figure out how many babies with a particular name are male.

To do this, we'll need to make a two variables to count the number of `male`s and `female`s of each name each year. We'll use the `ifelse()` function for each:

1. Make a `male` variable where, for each name in each year, if `sex=="M"`, then count the number of males (`n`) that year, otherwise set it equal to `0`.
2. Make a `female` variable where, for each name in each year, if `sex=="F"`, then count the number of females (`n`) that year, otherwise set it equal to `0`.

```{r}
babynames %>%
  mutate(male = ifelse(sex == "M", n, 0),
         female = ifelse(sex == "F", n, 0))
```

Now with this variable, we want to count the total number of males and females with each name over the entire dataset. Let's first `group_by(name)` so we'll get one row for every name. We will `summarize()` and take the `sum` of our `male` and of our `female` variables.

```{r}
babynames %>%
  mutate(male = ifelse(sex == "M", n, 0),
         female = ifelse(sex == "F", n, 0)) %>%
  group_by(name) %>%
    summarize(Male = sum(male),
              Female = sum(female))
```

Now, we want to figure out what *fraction* of each name is Male or Female. It doesn't matter which we do here, I'll do Male. `mutate()` a new variable I'll call `perc_male` for the percent of the name being for Male babies. It takes the summed variables we made before, and takes the fraction that are Male, multiplying by 100 to get percents (which isn't necessary, but is easy to read).

```{r}
babynames %>%
  mutate(male = ifelse(sex == "M", n, 0),
         female = ifelse(sex == "F", n, 0)) %>%
  group_by(name) %>%
    summarize(Male = sum(male),
              Female = sum(female))%>%
  filter(Male!=0,
         Female!=0) %>%
  mutate(perc_male = (Male/(Male+Female)*100))
```

Right now, it's still in alphabetical order. We want to arrange it by `perc_male`, and more importantly, we want `perc_male` to be between 48 and 52, so let's `filter` accordingly:

```{r}
babynames %>%
  mutate(male = ifelse(sex == "M", n, 0),
         female = ifelse(sex == "F", n, 0)) %>%
  group_by(name) %>%
    summarize(Male = sum(male),
              Female = sum(female))%>%
  mutate(perc_male = (Male/(Male+Female)*100)) %>%
  arrange(perc_male) %>%
  filter(perc_male > 48,
         perc_male < 52)
```

This gives us a lot of names, all falling between 48% and 52% male. But we want the most popular names that are in this range. So let's finally `mutate` a new variable called `total` that simply adds the number of `Male` and `Female` babies with a name. Then let's `arrange` our results by `desc(total)` to get the largest first, and then `slice(1:10)` to get the top 10 only.

```{r}
babynames %>%
  mutate(male = ifelse(sex == "M", n, 0),
         female = ifelse(sex == "F", n, 0)) %>%
  group_by(name) %>%
    summarize(Male = sum(male),
              Female = sum(female))%>%
  mutate(perc_male = (Male/(Male+Female)*100)) %>%
  arrange(perc_male) %>%
  filter(perc_male > 48,
         perc_male < 52) %>%
  mutate(total = Male+Female) %>%
  arrange(desc(total)) %>%
  slice(1:10)
```

@@

I am creating this document for/from the website, so these are all stored in a folder called `data`, one folder up from my current folder, `homeworks`. To get there, I go up one folder (`..`) and move to `data`, where these `csv` files are stored.

I suggest you either keep the data in the same folder as your `R` working directory (always check with `getwd()`), or create an R Project and store the data files in that same folder.

```{r}
# import data with read_csv from readr

# note these file paths will be different for you
polfreedom<-read_csv("../data/freedomhouse2018.csv")
econfreedom<-read_csv("../data/econfreedom.csv")

# look at each dataframe
polfreedom
econfreedom
```

## 9. 
**The `polfreedom` dataset is still a bit messy. Let's overwrite it (or assign to something like `polfreedom2`) and select `Country/Territory` and `Total` (total freedom score) and rename `Country.Territory` to `Country`.**

```{r}
polfreedom<-polfreedom %>%
  select(`Country/Territory`, Total) %>%
  rename(Country=`Country/Territory`)
```

@@

This one is a bit advanced to explain (but see the last few slides of 1.5 for more), so just copy what I gave you!

@@

```{r, fig.retina=3}
# install.packages("ggrepel") install for first use 
library(ggrepel) # load

### 2019-09-25 static/slides/09-slides.Rmd
Add problem set 1 answers, add broom logo in slides 2.3

@@

# tidy regression output
tidy(school_reg)
```
]

- The `broom` package allows us to tidy up regression objects

- The `tidy()` function creates a tidy tibble of regression output
]

@@

```{r, echo=T, eval=F}
# load packages
library(broom)

# tidy regression output
tidy(school_reg,
     conf.int = TRUE) # add CI's #<<
```

```{r, echo=F}
tidy(school_reg, conf.int = TRUE)
```
]

- The `broom` package allows us to tidy up regression objects

- The `tidy()` function creates a tidy tibble of regression output

- Add `conf.int = TRUE` option to include confidence intervals for each coefficient
]

>>>>

.left-column[
.center[
![](https://www.dropbox.com/s/qpe1604cf7s9r0g/rbroom.png?raw=1)
]

]
.right-column[

- The `broom` package allows us to *tidy* up regression objects<sup>.red[1]</sup>

- The `tidy()` function creates a *tidy* `tibble` of regression output

@@

# tidy regression output
tidy(school_reg)
```

.footnote[<sup>.red[1]</sup> See more at [broom.tidyverse.org](https://broom.tidyverse.org/).]
---

@@

.left-column[
.center[
![](https://www.dropbox.com/s/qpe1604cf7s9r0g/rbroom.png?raw=1)
]

]
.right-column[

- The `broom` package allows us to *tidy* up regression objects<sup>.red[1]</sup>

- The `tidy()` function creates a *tidy* `tibble` of regression output

# tidy regression output (with confidence intervals!)
tidy(school_reg,
     conf.int = TRUE) #<<
```

.footnote[<sup>.red[1]</sup> See more at [broom.tidyverse.org](https://broom.tidyverse.org/).]
